### 2021-07-20 presentations/presentation_current/presentation.tex
renamed stuff will make a new presentation

% Created 2021-07-20 Tue 18:30
% Intended LaTeX compiler: pdflatex
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{natbib}
\renewcommand{\bibsection}{}
\usetheme{metropolis}
\author{Benjamin Goldman}
\date{\today}
\title{Presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\begin{frame}[label={sec:org6b28ed4}]{Test}
\end{frame}
\end{document}