### 2021-07-21 presentations/presentation_current/presentation.tex
new r stuff outlined presentation yay

% Created 2021-07-20 Tue 20:15
% Intended LaTeX compiler: pdflatex
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{natbib}
\renewcommand{\bibsection}{}
\usetheme{default}
\usecolortheme{owl}
\author{Benjamin Goldman}
\date{\today}
\title{Presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Introduction}
\label{sec:orgc44fc10}

\begin{frame}[label={sec:orgf174f87}]{El Nino}
\end{frame}
\begin{frame}[label={sec:orgbb4e216}]{Climate Change}
\end{frame}
\begin{frame}[label={sec:org9a7241e}]{Chaos/Timeseries analysis}
\end{frame}
\begin{frame}[label={sec:org0558743}]{Gap/rationale}
\end{frame}
\begin{frame}[label={sec:orgaa26d89}]{Hypothesis}
\end{frame}

\section{Materials, Methods, and Results}
\label{sec:org45d9343}
\begin{frame}[label={sec:orgfdd47ac}]{Climate simulation}
\end{frame}
\begin{frame}[label={sec:org59b9f58}]{CESM1 and CESM2}
\end{frame}
\begin{frame}[label={sec:org8e1af0f}]{R and Python tools}
**
\end{frame}

\section{Discussion and Conclusion}
\label{sec:orgc16fbba}
\end{document}