### 2021-07-22 presentations/presentation_current/presentation.tex
presentation skeleton

\documentclass{beamer}
\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
\author{Ben Goldman}
\date{\today}
\usepackage{natbib}
\usepackage{tikz}
\usetheme{metropolis}
\renewcommand{\bibsection}{}

\begin{document}

\maketitle

\begin{frame}
  \frametitle{Climate Change}
  \citet{bjerknes1969atmospheric}

\end{frame}

\begin{frame}
  \frametitle{El Niño}

\end{frame}

\begin{frame}
  \frametitle{Climate Simulation}

\end{frame}

\begin{frame}
  \frametitle{ENSO in the Future}

\end{frame}

\begin{frame}
  \frametitle{Gap and Goal}

\end{frame}

\begin{frame}
  \frametitle{Research Questions}

\end{frame}

\begin{frame}
  \frametitle{signal and Noise}

\end{frame}

\begin{frame}
  \frametitle{Ensembles: CESM1 and CESM2}

\end{frame}

\begin{frame}
  \frametitle{R and Python tools}

\end{frame}

\begin{frame}
  \frametitle{Role of Mentor and Student}

\end{frame}

\begin{frame}
  \frametitle{Measuring ENSO}

\end{frame}

\begin{frame}
  \frametitle{Measuring ENSO Intensity}

\end{frame}

\begin{frame}
  \frametitle{ENSO is Becoming Stronger}

\end{frame}

\begin{frame}
  \frametitle{It's not That Simple}

\end{frame}

\begin{frame}
  \frametitle{Single Forcing Ensembles}

\end{frame}

\begin{frame}
  \frametitle{Influence of Aerosols and Greenhouse Gasses}

\end{frame}

\begin{frame}
  \frametitle{Correlation With Ocean Temperature}

\end{frame}

\begin{frame}
  \frametitle{Stratification}

\end{frame}

\begin{frame}
  \frametitle{Stratification in Other Ensembles}

\end{frame}

\begin{frame}
  \frametitle{References}
  \bibliographystyle{apalike}
  \bibliography{references.bib}
\end{frame}

\maketitle

\end{document}