### 2021-08-09 paperwork/paper/paper.tex

% Created 2021-08-08 Sun 15:05
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[letterpaper, margins=1in]{geometry}
\author{Benjamin Goldman}
\date{\today}
\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={The Impact of Anthropogenic Forcing on ENSO Amplitude},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:org1b56736}
\section{Data}
\label{sec:orgaaab045}
\section{Methods and Results}
\label{sec:org3f97e45}
\section{Conclusion}
\label{sec:org0e6ec22}
\section{Acknowledgments}
\label{sec:orgaaa58bc}

\bibliographystyle{apalike}
\bibliography{references}
\end{document}