### 2021-10-22 feedback/presentation_main.tex

% Created 2021-10-19 Tue 07:57
% Intended LaTeX compiler: pdflatex
\documentclass[little]{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\setcounter{secnumdepth}{1}
\author{Benjamin Goldman}
\date{\today}
\title{Feedback for main presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Feedback for main presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Feedback 1}
\label{sec:org1da6f23}
Date: \textit{<2021-09-26 Sun>}
Time: 20:20
Listener: Abby Goldman
\subsection*{Feedback}
\label{sec:org1d74604}
\begin{itemize}
\item lots of arm waving. SOOOOMUCH ARM WAVING!!! And very little breathing
\item make a list of “overhead” - make sure all concepts are well defined, in your notes, possibly in your slides
\item summary of what your many slides are — many different crunching of different data to circle around a process that we’re trying to under
\item insert heading slides, overall need a good deal more organization.
\item slide 1: break up into two or even 3?
\item Slide 2: Need to explain more generally what’s in the figure
\item Slide 3: climate simulation — is this method? If so, say so
\item Slide 4: ENSO and climate change slide — have “literature review” in heading.
\item slide about Mentor: do a better job explaining the interactive process between you and your mentor.
\item 3.4 index - whahah? I didn’t follow
\item Single forcing ensembles slide — not sure if all those arrows are working for you. Skip over single forcing setup, maybe?
\item use consistent language about the aerosols and their influence, considering that it’s a prediction
\end{itemize}

\section{Feedback 2}
\label{sec:org11974c6}
Date: \textit{<2021-09-30 Thu>}
Time: 8:20
Listener: Marcus Chung, Sean O'Donnell, Jimena Perez
\subsection*{Feedback:}
\label{sec:orgdbbe7e5}
\begin{itemize}
\item Figure text is too small
\item Caption text can be smaller to make room
\item Say some transition for the conclusion heading slide
\item Clarify difference between El Niño and La Niña
\begin{itemize}
\item overall make intro 2 more organized
\end{itemize}
\item Good understanding and well communicated
\item Are some slides too empty-looking?
\item Say fewer words because it is over time and talking too fast
\item Use note cards
\item Periods should be consistent.
\end{itemize}

\section{Feedback 3}
\label{sec:org52baac0}
Date: \textit{<2021-09-30 Thu>}
Time: 16:00
Listener: Michael Goldman
\subsection*{Feedback:}
\label{sec:orgf282c69}
\begin{itemize}
\item Make a few bullets clearer
\item Good tone
\item Slow down pacing
\item Introduction is slightly unclear
\item Use note cards
\end{itemize}

\section{Feedback 4}
\label{sec:org585437d}
Date: \textit{<2021-10-17 Sun>}
Time: 17:12
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgeffe568}
\begin{itemize}
\item Make figures bigger and make captions smaller
\item Love how detailed the content is
\item Great application slide
\item Make sure you explain your results in the context of your study, think big picture
\item Intro slide 2 is great we’ll explained
\end{itemize}
\end{document}