### 2021-12-15 presentations/board/middle.tex

\documentclass{beamer}

\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
\author{Ben Goldman}
\date{\today}

\usepackage{natbib}
\usepackage{tikz}
\usepackage{varwidth}
\usepackage{hyperref}
\usepackage[scale=1.4]{beamerposter}
\geometry{papersize={32in,48in}}
\usepackage{lipsum}

\usetikzlibrary{arrows,snakes,backgrounds}
\usetikzlibrary{positioning}
\tikzstyle{process}=[rectangle, draw=process, fill=process!20, line width = 0.3mm]
\tikzstyle{data}=[rectangle, draw=data, fill=data!20, line width = 0.3mm]
\tikzstyle{tight}=[node distance = 0.5in]
\tikzstyle{loose}=[node distance = 2in]
\definecolor{process}{HTML}{3d5f8f}
\definecolor{data}{HTML}{8f6d3d}

\usetheme{ben}

\newcommand{\myfig}[4]{
  \begin{figure}
    \centering
    \includegraphics[width=#3\textwidth]{figures/#1}
    \caption{#2}
    \label{fig:#4}
  \end{figure}
}

\renewcommand{\bibsection}{}

\begin{document}
\begin{frame}{The Impact of Anthropogenic Forcing on ENSO Amplitude}{Ben Goldman}

  \begin{block}{ENSO is Becoming Stronger (Past year results)}
    \begin{columns}
      \column{0.4\textwidth}
      \begin{itemize}
      \item Increase in ENSO intensity in both ensembles.
      \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
      \end{itemize}
      \column{0.5\textwidth}
      \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{1.0}{ff_compare}
    \end{columns}
  \end{block}

  \vfill

  \begin{block}{Influence of Aerosols and Greenhouse Gasses (Past year results)}
    \begin{columns}
      \column{0.4\textwidth}
      \begin{itemize}
      \item Influence of each factor on ENSO amplitude.
      \item Increased variance due to greenhouse gas emissions.
      \item Somewhat increased variance from aerosol emissions, but not linear.
      \end{itemize}
      \alert{Takeaway:} Human activities are triggering predicted strengthening of ENSO.
      \column{0.5\textwidth}
      \myfig{cesm1_sf_4.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{1.0}{cesm1_sf_4}
    \end{columns}
  \end{block}

  \vfill

  \begin{block}{Correlation With Ocean Temperature (Current results)}
    \begin{columns}
      \column{0.4\textwidth}
      \begin{itemize}
      \item Correlation coefficient between ocean temperature and ENSO amplitude.
      \item Negative coefficient in subsurface layer.
      \item Positive coefficient in surface layer.
      \item Suggests that ocean stratification may be mediating global warming influence on ENSO.
      \item Difference in heating modifies mechanics of ENSO cycle.
      \end{itemize}
      \column{0.5\textwidth}
      \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{1.0}{tempdt}
    \end{columns}
  \end{block}

  \vfill

\end{frame}
\end{document}